Bayesian determination of the effect of a deep eutectic solvent on the structure of lipid monolayers

### 2018-11-06 reports/paper.tex
toms affiliation ordered numerically
move usage statement to references

\documentclass[amsmath,amssymb,twocolumn,superscriptaddress,prl]{revtex4-1}

@@

\author{T.~Arnold}
\email{tom.arnold@esss.se}
\affiliation{European Spallation Source, SE-211 00 Lund, Sweden}
\affiliation{ISIS Neutron and Muon Source, Science and Technology Facilities
Council, Rutherford Appleton Laboratory, Harwell Oxford, Didcot, OX11 0QX,
UK}
\affiliation{Diamond Light Source, Harwell Campus, Didcot, OX11 0DE, UK}
\affiliation{Department of Chemistry, University of Bath, Claverton Down,
Bath, BA2 7AY, UK}

@@

%\tableofcontents
\begin{description}
\item[Usage]
Electronic Supplementary Information (ESI) available: All analysis/plotting
scripts and figure files, allowing for a fully reproducible, and automated,
analysis workflow for the work presented is available at
\url{https://github.com/arm61/lipids_at_airdes} (DOI: 10.5281/zenodo.1464836)
under a CC-BY 4.0 license.
Reduced experimental datasets are available at
\url{https://researchdata.bath.ac.uk/id/eprint/548}, under a CC-BY 4.0 license.
\end{description}

Deep eutectic solvents (DES) are green, sustainable liquids that are obtained
through the combination of ionic species with compounds that act as hydrogen
bond donors, such as sugars, alcohols, amines, and carboxylic
acids\cite{Smith2014,Dai2013}.
The resulting extensive hydrogen bonding network is able to stabilise the
ionic species and allows the eutectic mixture to remain liquid at room
temperature\cite{Hammond2016,Hammond2017,Araujo2017}.
Through different combinations of the precursor materials, it is possible to
tune the solvent's physicochemical properties, such as
polarity\cite{Pandey2014}, viscosity and surface tension\cite{Smith2014},
network charge\cite{Zahn2016}, and
hydrophobicity\cite{Ribeiro2015,vanOsch2015}.
Recently DES have also been shown to exhibit a ``solvophobic'' effect through
the promotion of surfactant micelle
formation\cite{Sanchez-Fernandez2016,Arnold2015,Hsieh2018,Banjare2018},
phospholipid bilayer formation\cite{Bryant2017,Bryant2016,Gutierrez2009}, and
the ability to stabilise non-ionic polymer\cite{Sapir2016} and protein
conformations\cite{Sanchez-Fernandez2017}.

@@

%
\begin{figure}[b]
\includegraphics[width=0.3\textwidth]{figures/head_groups}
\caption{\label{fig:heads}\small The two lipid classes with different head
groups compared in this study, where R indicates the hydrocarbon tail;
(a) phosphatidylglycerol (PG), (b) phosphocholine (PC).}
\end{figure}
%
We report on measurements of four phospholipids monolayers, namely
1,2-dipalmitoyl-sn-glycero-3-phosphocholine (DPPC, C$_{16}$ tails),
1,2-dimyristoyl-sn-glycero-3-phosphocholine (DMPC, C$_{14}$ tails),
1,2-dilauroyl-sn-glycero-3-phosphocholine (DLPC, C$_{12}$ tails) and
1,2-dimyristoyl-sn-glycero-3-phospho-(1'-rac-glycerol)
(DMPG, C$_{14}$ tails), at the air-DES (1:2 choline choride:glycerol)
interface.
In contrast to many previous studies\cite{Mohwald1990,Kewalramani2010,
Bayerl1990,Johnson1991,Clifton2012,Helm1987,Daillant1990}, we have developed
a chemically-consistent model (detail in the ESI) that allows for the
co-refinement of reflectometry measurements at different surface pressure
and makes no assumption of the volume of the lipid head, $V_h$, or tail, $V_t$.
Instead these parameters were allowed to vary for each lipid while being
constrained to be self-consistent over different surface pressures in the
same phase; Liquid-Condensed (LC) for DPPC and Liquid-Expanded (LE) for
DMPC, DMPG and DLPC.
This model was required because we cannot assume that the electrostatic
interactions between lipid head and the solvent are the same in the DES and
in water. This may therefore influence the effective head volume which means
we cannot rely on the literature values (see ESI) derived from measurements
on or in water.
Furthermore, it is known that, on water, increased surface pressure and
associated LE-LC phase transitions lead to a compression of the lipid tail
volume\cite{Marsh2010,Small1984} and this compaction has not necessarily
been accounted for in the literature\cite{Campbell2018}.
Our approach avoids this issue by making no assumption about the molecular
volumes and only considering surface pressures that we believe to be in the
same phase.

>>>>

\documentclass[amsmath,amssymb,twocolumn,superscriptaddress,aps,prl]{revtex4-1}

@@

\author{T.~Arnold}
\email{tom.arnold@esss.se}
\affiliation{Department of Chemistry, University of Bath, Claverton Down,
Bath, BA2 7AY, UK}
\affiliation{European Spallation Source, SE-211 00 Lund, Sweden}
\affiliation{Diamond Light Source, Harwell Campus, Didcot, OX11 0DE, UK}
\affiliation{ISIS Neutron and Muon Source, Science and Technology Facilities
Council, Rutherford Appleton Laboratory, Harwell Oxford, Didcot, OX11 0QX,
UK}

@@

%\tableofcontents
Deep eutectic solvents (DES) are green, sustainable liquids that are obtained
through the combination of ionic species with compounds that act as hydrogen
bond donors, such as sugars, alcohols, amines, and carboxylic
acids\cite{Smith2014,Dai2013}.
The resulting extensive hydrogen bonding network is able to stabilise the
ionic species and allows the eutectic mixture to remain liquid at room
temperature\cite{Hammond2016,Hammond2017,Araujo2017}.
Through different combinations of the precursor materials, it is possible to
tune the solvent's physicochemical properties, such as
polarity\cite{Pandey2014}, viscosity and surface tension\cite{Smith2014},
network charge\cite{Zahn2016}, and
hydrophobicity\cite{Ribeiro2015,vanOsch2015}.
Recently DES have also been shown to exhibit a ``solvophobic'' effect through
the promotion of surfactant micelle
formation\cite{Sanchez-Fernandez2016,Arnold2015,Hsieh2018,Banjare2018},
phospholipid bilayer formation\cite{Bryant2017,Bryant2016,Gutierrez2009}, and
the ability to stabilise non-ionic polymer\cite{Sapir2016} and protein
conformations\cite{Sanchez-Fernandez2017}.

@@

%
\begin{figure}[b]
\includegraphics[width=0.3\textwidth]{figures/head_groups}
\caption{\label{fig:heads}\small The two lipid classes with different head
groups compared in this study, where R indicates the hydrocarbon tail;
(a) phosphatidylglycerol (PG), (b) phosphocholine (PC).}
\end{figure}
%
We report on measurements of four phospholipids monolayers, namely
1,2-dipalmitoyl-sn-glycero-3-phosphocholine (DPPC, C$_{16}$ tails),
1,2-dimyristoyl-sn-glycero-3-phosphocholine (DMPC, C$_{14}$ tails),
1,2-dilauroyl-sn-glycero-3-phosphocholine (DLPC, C$_{12}$ tails) and
1,2-dimyristoyl-sn-glycero-3-phospho-(1'-rac-glycerol)
(DMPG, C$_{14}$ tails), at the air-DES (1:2 choline choride:glycerol)
interface.
In contrast to many previous studies\cite{Mohwald1990,Kewalramani2010,
Bayerl1990,Johnson1991,Clifton2012,Helm1987,Daillant1990}, we have developed
a chemically-consistent model that allows for the
co-refinement of reflectometry measurements at different surface pressure
and makes no assumption of the volume of the lipid head, $V_h$, or tail, $V_t$ \cite{SI2018}.
Instead these parameters were allowed to vary for each lipid while being
constrained to be self-consistent over different surface pressures in the
same phase; Liquid-Condensed (LC) for DPPC and Liquid-Expanded (LE) for
DMPC, DMPG and DLPC.
This model was required because we cannot assume that the electrostatic
interactions between lipid head and the solvent are the same in the DES and
in water. This may therefore influence the effective head volume which means
we cannot rely on the literature values derived from measurements
on or in water \cite{SI2018}.
Furthermore, it is known that, on water, increased surface pressure and
associated LE-LC phase transitions lead to a compression of the lipid tail
volume\cite{Marsh2010,Small1984} and this compaction has not necessarily
been accounted for in the literature\cite{Campbell2018}.
Our approach avoids this issue by making no assumption about the molecular
volumes and only considering surface pressures that we believe to be in the
same phase.
